---
title: "p8105_hw2_xl2834"
author: "Xiaoyu Li"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr.TrashWheel dataset.
```{r}
trash_df = 
  read_xlsx(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data
```{r}
precip_2018 = 
   read_xlsx(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
     sheet = "2018 Precipitation",
     skip = 1
  ) %>%
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
   read_xlsx(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
     sheet = "2017 Precipitation",
     skip = 1
  ) %>%
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

```


Combine annual precipitation
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```

The dataset trash_df contains information about Mr. Trashwheel trash collector in Baltimore, Maryland. The dataset include variables `r ls(trash_df)`. 
There are a total of `r nrow(trash_df)` observations in the dataset. The total weight of trash collected is `r sum(trash_df$weight_tons)` tons. The median number of sports balls in a dumpster in 2017 is `r trash_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.


The dataset precip_df contains information about precipitation in Baltimore, Maryland in 2017 and 2018. The dataset include variables `r ls(precip_df)`.  The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. There are a total of `r nrow(trash_df)` observations in the dataset.


## Problem 2

```{r}
subway_df = read_csv(
  "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "yes", TRUE, FALSE))
```

The dataset contains information about subway station in NYC. The dataset include variables `r ls(subway_df)`. I have made the variable names clearer, kept some useful variables and converted a character variable to a logical variable. The dimension of the dataset is 'dim(subway_df)`. The data are still not tidy.


```{r}

```


There are `r nrow(distinct(subway_df, line, station_name))` distinct stations.

```{r}
newdf = 
  distinct(subway_df, line, station_name, .keep_all = TRUE)
count(newdf)
```


```{r}
newdf %>%
  filter(ada == TRUE) %>% 
  nrow()
```

#### What proportion of station entrances / exits without vending allow entrance?
```{r}
subway_df %>% 
  filter(vending == "NO") %>% 
  count()/count(subway_df)
```

The proportion of station entrances/exits without vending is 9.8%.


```{r}
subway_df %>% 
  mutate(route8:route11 = as.character(route8:route11))

subway_tidy = 
  subway_df%>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    values_to = "name"
    
  )
```

