---
title: "p8105_hw2_xl2834"
author: "Xiaoyu Li"
date: "9/24/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr.TrashWheel dataset.
```{r}
trash_df = 
  read_xlsx(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data
```{r}
precip_2018 = 
   read_xlsx(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
     sheet = "2018 Precipitation",
     skip = 1
  ) %>%
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
   read_xlsx(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
     sheet = "2017 Precipitation",
     skip = 1
  ) %>%
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

```


Combine annual precipitation
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```

The dataset trash_df contains information about Mr. Trashwheel trash collector in Baltimore, Maryland. The dataset include variables `r ls(trash_df)`. 
There are a total of `r nrow(trash_df)` observations in the dataset. The total weight of trash collected is `r sum(trash_df$weight_tons)` tons. The median number of sports balls in a dumpster in 2017 is `r trash_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.


The dataset precip_df contains information about precipitation in Baltimore, Maryland in 2017 and 2018. The dataset include variables `r ls(precip_df)`.  The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. There are a total of `r nrow(trash_df)` observations in the dataset.


## Problem 2

```{r}
subway_df = read_csv(
  "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude,station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))

```

The dataset contains information about subway station in NYC. The dataset include variables `r ls(subway_df)`. I have made the variable names clearer, kept some useful variables and converted a character variable to a logical variable. The dimension of the dataset is `dim(subway_df)`. The data are still not tidy, because the route information is spreading over 11 columns.

#### How many distinct stations are there?

```{r}
subway_df %>% 
  distinct(station_name, line) %>% 
  nrow()
```
There are 465 distinct stations.

#### How many stations are ADA compliant?

```{r}
subway_df %>%
  distinct(station_name, line, .keep_all = TRUE) %>%
  filter(ada == TRUE) %>% 
  nrow()
```
84 are ADA compliant.

#### What proportion of station entrances / exits without vending allow entrance?
```{r}
no_vending = 
  filter(subway_df, vending == "NO")

no_vending %>% 
  filter(entry == TRUE) %>% 
  count()/count(no_vending)
```

Among the station entrance/exits without vending, there are 37.7% that allow entrance.


#### Reformat the data so that route number and route name are distinct variables.

```{r}
subway_df = 
  mutate(subway_df,
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
         )

subway_tidy = 
  subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    names_prefix = "route",
    values_to = "route_number"
    )
```

#### How many distinct stations serve the A train?
```{r}
subway_tidy %>% 
  filter(route_number == "A") %>% 
  distinct(station_name, line) %>% 
  nrow()
```
There are 60.

#### Of the stations that serve the A train, how many are ADA compliant?

```{r}
subway_tidy %>% 
  filter(route_number == "A", ada == TRUE) %>% 
  distinct(station_name, line) %>%
  nrow()
```

There are 17.

## Problem 3

#### Step 1 Clean the data in pols-month.csv.
```{r}
pols_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  pivot_longer(c(prez_gop, prez_dem),
               names_to = "president",
               values_to = "value",
               names_prefix = "prez_"
              ) %>% 
  filter(value != 0) %>% 
  select(-day, -value) %>% 
  mutate(month = as.integer(month),
         year = as.integer(year)) %>% 
  mutate(month = month.abb[month])
  
```


#### Step 2 Clean the data in snp.csv using a similar process.
```{r}
snp_df = read_csv(
  "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  select(-day) %>% 
  relocate(year, month) %>% 
  mutate(month = as.integer(month),
         year = as.integer(year)) %>% 
  mutate(month = month.abb[month])
  
```

#### Step 3 Tidy the unemployment data.
```{r}
unemployment_df = read_csv(
  "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percentage"
  ) %>% 
    mutate(year = as.integer(year))
```

#### Join the datasets
```{r}
joined_df1 = 
  left_join(pols_df, snp_df, by = c("year", "month"))

unemployment_df = 
  mutate(unemployment_df, year = as.integer(year))

result_df =
  left_join(joined_df1, unemployment_df, by = c("year", "month"))
```


The dataset pols_df has `r ncol(pols_df)` variables and `r nrow(pols_df)` observations, and contains information related to the number of national politicians who are democratic or republican at any given time. There are variables showing the number of governors, senators and representatives for each party, and a variable indicating the party of the president at the given time. 

The dataset snp_df has `r ncol(snp_df)` variables and `r nrow(snp_df)` observations, and contains information about Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole.

The dataset unemployment_df has `r ncol(unemployment_df)` variables and `r nrow(unemployment_df)` observations, and shows the unemployment percentages in a given month in a year.

The result dataset has the dimension of `r dim(result_df)`. It includes information about number of governors, senators and representatives in each party, unemployment percentages, and Standard & Poor’s stock market index in year ranging from `r first(result_df$year)` to `r last(result_df$year)`.


